---
title: "Prediction Writeup"
author: "Stella"
date: '2022-12-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis
In this project, we use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.

## Load libraries

```{r}
library(caret)
library(randomForest)
```

## Load Data and delete NA

```{r}
training.url <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
test.cases.url <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
downloadcsv <- function(url, nastrings) {
    temp <- tempfile()
    download.file(url, temp, method = "curl")
    data <- read.csv(temp, na.strings = nastrings)
    unlink(temp)
    return(data)
}
train <- downloadcsv(training.url, c("", "NA", "#DIV/0!"))
test <- downloadcsv(test.cases.url, c("", "NA", "#DIV/0!"))
dim(train)
table(train$classe)
```

##Validation and Selection

```{r}
set.seed(12345)
trainset <- createDataPartition(train$classe, p = 0.8, list = FALSE)
Training <- train[trainset, ]
Validation <- train[-trainset, ]
nonzerocol <- nearZeroVar(Training)
Training <- Training[, -nonzerocol]
countlength <- sapply(Training, function(x) {
    sum(!(is.na(x) | x == ""))
})
nullCol <- names(countlength[countlength < 0.6 * length(Training$classe)])
descriptcol <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
    "cvtd_timestamp", "new_window", "num_window")
excludecolumns <- c(descriptcol, nullCol)
Training <- Training[, !names(Training) %in% excludecolumns]
```


## Creating models

```{r}
rfModel <- randomForest(as.factor(classe)~ ., data = Training, importance = TRUE, ntrees = 10)
ptraining <- predict(rfModel, Training)
```

## Validation and Test the model

```{r}
u1 <- union(ptraining,Training$classe)
t1 <- table(factor(ptraining, u1), factor(Training$classe, u1))
print(confusionMatrix(t1))
pvalidation <- predict(rfModel, Validation)
u2 <- union(pvalidation,Validation$classe)
t2 <- table(factor(pvalidation, u2), factor(Validation$classe, u2))
print(confusionMatrix(t2))
ptest <- predict(rfModel, test)
ptest
```



## Conclusions

Taking into account the previous data and the models; the result with the random forest model with cross validation, gives us an accurate result of approximately 99%.
